MCC Models

### 2019-09-03 papers/full_paper/manuscript.Rmd
migrate manuscript and complete SI

---
title             : "Predicting information integration in pragmatic word learning across development"

author: 
  - name          : "Manuel Bohn"
    affiliation   : "1,2"
    corresponding : yes    # Define only one corresponding author
    address       : "Department of Psychology, 450 Serra Mall, Stanford, CA 94305"
    email         : "bohn@stanford.edu"
  - name          : "Michael Henry Tessler"
    affiliation   : "3"
  - name          : "Megan Merrick"
    affiliation   : "1"
  - name          : "Michael C. Frank"
    affiliation   : "1"

@@

author_note: |
  MB received funding from the European Union’s Horizon 2020 research and innovation programme under the Marie Sklodowska-Curie grant agreement no. 749229. MCF was supported by a Jacobs Foundation Advanced Research Fellowship and the Zhou Fund. We thank Jacqueline Quirke and Sabina Zacco for help with the data collection.

abstract: |
  Language is used and learned in complex social settings. To infer what a speaker means, listeners have to go beyond the literal meaning of words and rely on pragmatic inferences. During language acquisition, the same processes can help children to learn the meaning of novel words. Yet, pragmatic inferences require balancing multiple sources of social information. Here we study this integration process. We isolate two types of pragmatic inference and formalize how they should be integrated according to a Bayesian pragmatics model. We present a series of experiments with preschool children and adults that suggest that information integration is best described as a form of probabilistic inference.

@@

@@

```{r}
# preparing datafiles (delete eventually)

# d <- read_csv(file="../../stats/data/inf.data.csv") %>%
#   mutate(trial_type = ifelse(trial == "train", "train", "test"),
#          trial = alltrial -2,
#          experiment = "adult_ex1",
#          condition = ifelse(control == TRUE, "control", "test") )%>%
#   filter(trial_type == "test")%>%
#   select(experiment, id,  trial,condition, agent, correct)
# 
# write_csv(d , path="../../stats/data/adult_ex1.csv")
# 
# d <- read_csv(file="../../stats/data/novel.data.csv") %>%
#   mutate(trial_type = ifelse(trial == "train", "train", "test"),
#          experiment = "adult_ex2_novelty",
#          speaker = ifelse(change == TRUE, "different_speaker" , "same_speaker"),
#          agent = ifelse( condition == "same_speaker", agent, altAgent))%>%
#   filter(trial != "train")%>%
#   select(experiment, id, trial,speaker, agent, correct)
# 
# write_csv(d , path="../../stats/data/adult_ex2_novelty.csv")
# 
# d <- read_csv(file="../../stats/data/pref.data.csv") %>%
#   mutate(trial_type = ifelse(trial == "train", "train", "test"),
#          experiment = "adult_ex2_preference",
#          speaker = ifelse(change == TRUE, "different_speaker" , "same_speaker"),
#          agent = ifelse( condition == "same_speaker", agent, altAgent))%>%
#   filter(trial != "train")%>%
#   select(experiment, id, trial,speaker, agent, correct)
# 
# write_csv(d , path="../../stats/data/adult_ex2_preference.csv")
# 
# d <- bind_rows(read_csv(file="../../stats/data/ex3.novel.data.csv"),
#                read_csv(file="../../stats/data/ex3.pref.data.csv")) %>%
#   distinct(id, alltrial, .keep_all = TRUE) %>%
#   mutate(trial_type = ifelse(trial == "train1" | trial =="train2", "train", "test"))
# 
# exclude <- d %>%
#   filter(trial_type == "train") %>%
#   group_by(id)%>%
#   summarise(correct_inf = mean(correct_inf)) %>%
#   filter(correct_inf == 0)
# 
# d <- d%>%
#   filter(!id %in% exclude$id)%>%
#   mutate(agent = ifelse( change == "same", agent, altAgent),
#          common_ground_manipulation = ifelse(experiment == "novel_inf","novelty", "preference"),
#          experiment = "adult_ex3",
#          trial = alltrial -2,
#          speaker = ifelse(change == "same", "same_speaker", "different_speaker"))%>%
#   filter(trial_type == "test")%>%
#   select(experiment, id,trial, common_ground_manipulation, speaker ,alignment, agent, correct_inf)
# 
# write_csv(d , path="../../stats/data/adult_ex3.csv")
# 
# 
#
# d <- bind_rows(read_csv(file="../../stats/data/ex3.2.novel.strong.data.csv"),
#                read_csv(file="../../stats/data/ex3.2.novel.medium.data.csv"),
#                read_csv(file="../../stats/data/ex3.2.novel.weak.data.csv"),
#                read_csv(file="../../stats/data/ex3.2.pref.strong.data.csv"),
#                read_csv(file="../../stats/data/ex3.2.pref.medium.data.csv")
#                ) %>%
#   distinct(id, alltrial, .keep_all = TRUE) %>%
#   mutate(trial_type = ifelse(trial == "train1" | trial =="train2", "train", "test"))
#          
# exclude <- d %>%
#   filter(trial_type == "train") %>%
#   group_by(id)%>%
#   summarise(correct_inf = mean(correct_inf)) %>%
#   filter(correct_inf == 0)
# 
# d <- d%>%
#   filter(!id %in% exclude$id)%>%
#   mutate(agent = ifelse( change == "same", agent, altAgent),
#          common_ground_manipulation = ifelse(grepl("novel",experiment),"novelty", "preference"),
#          trial = alltrial -2,
#          speaker = ifelse(change == "same", "same_speaker", "different_speaker"),
#          prior_manipulation = prior,
#          experiment = "adult_ex4")%>%
#   filter(trial_type == "test")%>%
#   select(experiment, id,trial, common_ground_manipulation, prior_manipulation, speaker, alignment, agent, correct_inf)
# 
# write_csv(d , path="../../stats/data/adult_ex4.csv")
#
# 
# d <- bind_rows(read_csv(file="../../stats/data/kids_pref_data.csv")%>%filter(age != "2")%>%mutate(speakerChange = change),
#                        read_csv(file="../../stats/data/kids_pref_2_data.csv")%>% mutate(age = subage))%>%
#   filter(trial != "train")%>%
#   mutate(age_bin = factor(age),
#          experiment = "child_ex2",
#          condition = ifelse(speakerChange == F, "same_speaker", "different_speaker"),
#          agent = ifelse( speakerChange == F, agent, altAgent),
#          id = subid,
#          minage = min(age_num))%>%
#   filter(trial != "train")%>%
#   select(experiment,id,age_bin,age_num,minage,trial,condition,agent,correct)
# 
# write_csv(d , path="../../stats/data/child_ex2.csv")
# 
# d <- bind_rows(
#   read_csv(file="../../stats/data/kids_inf_data.csv"),
#   read_csv(file="../../stats/data/kids_inf_2_data.csv"))%>%
#   filter(subage != "2",
#          trial != "filler1",
#          trial != "filler2")%>%
#   mutate(age_bin = factor(subage),
#          experiment = "child_ex1",
#          id =subid,
#          minage = min(age_num))%>%
#   select(experiment,id,age_bin,age_num,sex,minage,trial,correct)
# 
# write_csv(d , path="../../stats/data/child_ex1.csv")
#
# d <- read_csv(file="../../stats/data/kids_ex3_data.csv")%>%
#   mutate(experiment = "child_ex3",
#          id = subid,
#          age_bin = subage,
#          speaker = Speaker,
#          alignment = Alignment,
#          minage = min(age_num))%>%
#   filter(trial != "filler1",trial != "filler2")%>%
#   select(experiment, id,age_bin,age_num,minage,trial, speaker, alignment, agent, correct_inf)
# 
#  write_csv(d , path="../../stats/data/child_ex3.csv")

@@

@@

Language use and language learning require integrating and balancing different sources of information. And even though linguistic cues, such as conventional syntax and semantics, may provide strong evidence about the intended meaning of an utterance, additional pragmatic inferences are necessary arrive at a precise interpretation [@levinson2000presumptive; @sperber2001relevance]. A second route to meaning is even more important in the process of language learning. In this case, inferences based on the social embedding of the utterance are key to understanding its meaning and learning the words contained therein [@clark2009first; @tomasello2009constructing]. In this paper, we explore how different forms of social information are integrated during word learning across development. More specifically, we look at information integration as a form of probabilistic inference.

Pragmatic inferences rest on assumptions about how communication *should* proceed. Interlocutors mutually expect each other to communicate in a cooperative way and this mutual expectation serves as the basis for inferences that go beyond the literal meaning of an utterance [@grice1991studies; @sperber2001relevance; @tomasello2008origins]. Paul Grice [-@grice1991studies] famously summarized this notion in his *Cooperative Principle*, which reads: “Make your contribution such as is required, at the stage at which it occurs, by the accepted purpose or direction of the talk exchange in which you are engaged.” Two sets of expectations follow from this principle. On the one hand, there is the general expectation that one follows the principle and communicates in a relevant and informative way. On the other hand, there is the expectation that communication is further specific to the communicative partner, that is, that it is adjusted to the the common ground shared between speaker and listener [@bohn2018common; @clark1996using].

Importantly, these expectations can also support language development [@clark2009first; @tomasello2009constructing]. Starting in infancy, children generally expect adults' communicative acts to provide them with relevant information [e.g @behne2005one]. And preschool children learn novel words by assuming that speakers are informative [@frank2014inferring]. In the absence of any additional information about the potential referent objects or the speaker, children expected a novel word to be intended to refer to an object it describes in a unique way [see also @stiller2015ad]. Person specific common ground information supports communication in infancy [e.g. @bohn2018social; @saylor2011s] and facilitates word learning in slightly older children. For example, two-year-olds keep track of which object is new to a speaker to decide which object the speaker is referring to when they use a novel word [@akhtar1996role; @diesendruck2004two]. Furthermore, when three to five year old children learn about a speaker's preference, they later use this information to identify the meaning of an unknown word coming from the same speaker [@saylor2009preschoolers]. These studies suggest early emerging pragmatic competences. Yet, communicative interactions in naturalistic contexts have multiple layers and require balancing expectations [@clark1996using]. So how are different expectations integrated?

The Rational Speech Act (RSA) framework [@frank2012predicting; @goodman2016pragmatic] offers a formal framework for addressing this integration problem. RSA models are characterized by their recursive structure. Each agent in the recursion is modeled as a Bayesian reasoner; thus, information integration is treated as a process of probabilistic inference. The expectation that speakers communicate informatively is already encoded in the structure of the model: Speakers choose utterances that help listeners in disambiguating referents. Common ground information can be integrated in the model as a shared prior probability of referents in the context of the utterance. Thus, a natural locus for information integration within RSA models is the trade off between the prior probability of a referent and the likelihood of that referent given the current utterance.

Here we apply this rational, pragmatic account of communication to the study of information integration during word learning. We do so from a developmental perspective and include children, continuously sampled between three and five years of age, as well as adults. For both groups, we first test speaker-specific and common ground expectations independently. Based on this data, we generate model predictions about how expectations should be integrated. Finally, we compare these predictions to newly collected data and use model comparisons to test different hypothesis about how expectations are integrated.

```{r fig1, include = T, fig.align = "center", fig.cap = "Schematic experimental procedure with screenshots from the adult experiments. In all conditions, at test (bottom), the speaker ambiguously requested an object using a non-word (e.g. “dax”). Participants clicked on the object they thought the speaker referred to. Informativeness (Experiment 1, left) translated to making one object less frequent in context. Common ground (Experiment 2, middle) was manipulated by making one object prefered by or new to the speaker. Green plus signs represent utterances that expressed preference and red minus of dispreference (see main text for details). Experiment 3 (right) combined manipulations. When expressing e.g. preference for an object on a table with two objects (panel 3 from top), the respective object was temporarily enlarged. Condition for Experiment 3 shown here: preference - same speaker - incongruent.", out.width="450px"}
knitr::include_graphics("./figures/fig1.png")
```

# Adults

## Assessment of general and speaker specific expectations (Experiment 1 and 2)

As a first step, we assessed whether adults make general and common ground inferences when tested separately. In experiment 1, participants could infer which object a novel word referred to by assuming that the speaker communicated in an informative way. The speaker was located between two tables, one with two novel objects, A and B, and the other with only object A (Fig. 1A). When the speaker turned to the table with the two objects (A and B) and used an novel word to request one of them, participants could infer that the word referred to object B. This follows from the counter-factual inferences that, if the (informative) speaker would have wanted to refer to object A, they would have turned to the table with the single object, because this would have been the least ambiguous way to refer to this object. As a consequence, turning to the table with two objects most likely reflects an intention to refer to object B. In the control condition, both tables contained both objects and no inference could be made based on the speaker's behavior.

```{r fig2, include = T, fig.align = "center", fig.cap = "Results from experiment 1 and 2 for adults. For preference and novelty, control refers to a different speaker (see Fig. 1B). Transparent dots show data from individual participants, diamonds represent condition means, error bars are 95\\% CIs. Dashed line indicates performance expected by chance.", out.width="250px"}
knitr::include_graphics("./figures/fig2.png")
```

@@

# model
lm_ex1 <- glmer(correct ~ condition +
              (1 |id) , 
              data = adult_ex1_data, 
              family = binomial,control=glmerControl(optimizer="bobyqa",optCtrl=list(maxfun=2e5)))

@@

Participants made the inference consistent with the assumption that the speaker communicated informatively: They selected object B above chance in the test condition (t(`r ex1t%>%filter(condition == "test")%>%pull(df)`) = `r ex1t%>%filter(condition == "test")%>%pull(t_value)`, *p* `r ex1t%>%filter(condition == "test")%>%pull(p_value)`, see Fig. 2) and more often compared to the control condition (*$\beta$* = `r ex1_r%>%filter(term == "conditiontest")%>%pull(estimate)`, se = `r ex1_r%>%filter(term == "conditiontest")%>%pull(std.error)`, *p* `r ex1_r%>%filter(term == "conditiontest")%>%pull(p.value)`).

```{r ex2-results}
ex2tpref <- adult_ex2_preference_data %>%
  group_by(condition, id) %>%
  summarise(correct = mean(correct)) %>%
  summarise(correct = list(correct)) %>%
  group_by(condition) %>%
  mutate(df= t.test(unlist(correct), mu = 0.5)$parameter,
         t_value = t.test(unlist(correct), mu = 0.5)$statistic,
         p_value = t.test(unlist(correct), mu = 0.5)$p.value) %>%
  select(condition,df,t_value,p_value)%>%
  mutate(p_value = ifelse(p_value<.001,"< .001",as.character(paste("=",substr(round(p_value,3),2,5),sep=" "))),
         t_value = round(t_value,2))

@@

In experiment 2, we tested if participants use common ground information that is specific to a speaker to infer the referent of a novel word. In the preference manipulation (Fig. 1B . left), the speaker expressed preference for one of two objects. When the speaker later used a novel word to requested an object, participants could infer the referent by assuming that the speaker was talking about the object they liked. In a control condition, a different speaker, whose preferences were unknown, made the request. Adults selected the preferred object above chance when the same speaker made the request (t(`r ex2tpref%>%filter(condition == "same_speaker")%>%pull(df)`) = `r ex2tpref%>%filter(condition == "same_speaker")%>%pull(t_value)`, *p* `r ex2tpref%>%filter(condition == "same_speaker")%>%pull(p_value)`, see Fig. 2) and more so compared to when a different speaker asked for an object ($\beta$ = `r ex2pref_r%>%filter(term == "conditionsame_speaker")%>%pull(estimate)`, se = `r ex2pref_r%>%filter(term == "conditionsame_speaker")%>%pull(std.error)`, *p* `r ex2pref_r%>%filter(term == "conditionsame_speaker")%>%pull(p.value)`).

In the novelty manipulation (Fig 1B - right), the speaker encountered one object on one of the tables. Later, a second object appeared. The second object was therefore new to the speaker. When the same speaker later used a novel word to requested an object (in a slightly excited way), participants could infer that it was the new object they are referring to. In contrast, when a different speaker, to whom both objects were equally new, made a request, no inference could be made based on the speaker's state of knowledge. In line with this reasoning, participants selected the new object above chance with the same speaker (t(`r ex2tnov%>%filter(condition == "same_speaker")%>%pull(df)`) = `r ex2tnov%>%filter(condition == "same_speaker")%>%pull(t_value)`, *p* `r ex2tnov%>%filter(condition == "same_speaker")%>%pull(p_value)`, see Fig. 2), and more often compared to the different speaker ($\beta$ = `r ex2nov_r%>%filter(term == "conditionsame_speaker")%>%pull(estimate)`, se = `r ex2nov_r%>%filter(term == "conditionsame_speaker")%>%pull(std.error)`, *p* `r ex2nov_r%>%filter(term == "conditionsame_speaker")%>%pull(p.value)`). Taken together, experiment 1 and 2 showed that adults use general and speaker specific expectations to identify the referent of a novel word. In the following, we studied how these expectations are integrated when combined experimentally.

@@

We modeled the integration of general and common ground expectations as a process of probabilistic inference. We used the results of experiment 1 and 2 to inform key parameters in the model's architecture. Within the RSA framework, general expectations that speakers (or listeners) communicate in a cooperative and informative way are reflected in the model's recursive structure. When interpreting an utterance, the hypothetical listener imagines that it was generated by a pragmatic speaker whose goal is to produce the most informative utterance. The pragmatic speaker is modeled as reasoning about what the most informative utterance would be for a naive listener, who does not know the novel words. Importantly, the utterance in this case includes not just the novel word but also the turning to one of the tables. The conditional probability that the listener interprets an utterance as referring to a particular referent is defined as:

$$P_L(r_s|u)\propto P_S(u|r_s)P_S(r_s)$$

Here, the term $P_S(u|r_s)$ denotes the likelihood that the pragmatic speaker will produce a particular utterance $u$ to refer to a referent $r$. It is defined in terms of a utility function $U_S(u;s)$ consisting of the surprisal of $u$ for a naive listener, who interprets $u$ according its literal semantics:

$$P_S(u|r_s)\propto exp(\alpha U_S(u;s))$$

This utility is maximized in order to decide which utterance to use to communicate about a particular referent, reflecting the expectation that the speaker communicates in an informative way. The absolute strength of $P_S(u|r_s)$ depends on a scalar value, $\alpha$, which can be interpreted as an indicator of how rational the speaker is in choosing utterances (i.e. as $\alpha$ increases, the speaker is more likely to choose the most informative utterance). We used the data from experiment 1 to specify how rational participants thought speakers were in the current setup. That is, we inferred which value of $\alpha$ would generate model predictions (assuming equal prior probability for each object) that corresponded to the average proportion of correct responses measured in experiment 1.

We treated common ground expectations as changes in the prior probability that a particular referent is being referred to ($P_S(r_s)$ in the first equation). For example, if the speaker expresses preference for object A, and the same speaker later requests an object, then A has a higher probability of being the referent of a subsequent utterance, regardless of any additional pragmatic considerations. We used the proportions measured in experiment 2 to specify the probability distribution across referents for the different common ground manipulations (novelty or preference, same or different speaker).

With $\alpha$ and $P_S(r_s)$ fixed based on experiment 1 and 2, we generated model predictions for situations in which general and common ground expectations had to be integrated (Fig. 1C). Expectations could either be congruent or incongruent. If, for example, object A was the more informative object (unique to that table) and also the one the speaker expressed preference for, expectations were congruent. Next, the speaker who made the request could either be the same as the one in the previous interaction or a different one. Finally, common ground information could be manipulated in the form of preference or novelty. This resulted in a total of 2 (novelty or preference) x 2 (same or different speaker) x 2 (congruent or incongruent) = 8 conditions for which we generated model predictions. In order to capture that behavioral data is to some extend noisy, all models included a noise parameter. Noise parameters ranged between 0 and 1 and reflect the proportion of responses that are estimated (based on the data) to be random instead of in line with model predictions. Including the noise parameter greatly improved the model fit to the data (see SI Appendix for details). We did not preregister the inclusion of a noise parameter for experiment 3 but did so for all subsequent experiments.

```{r ex3 GLMM, cache=F}
# model 
lm_ex3 <- glmer(correct_inf ~ common_ground_manipulation*speaker*alignment + (alignment|id), 
              data = adult_ex3_data, family = binomial, 
              control=glmerControl(optimizer="bobyqa",optCtrl=list(maxfun=2e5)))

>>>>

---
title             : "Pragmatic cue integration in adults’ and children’s inferences about novel word meanings"
shorttitle        : "Pragmatic cue integration"

author: 
  - name          : "Manuel Bohn"
    affiliation   : "1,2"
    corresponding : yes    # Define only one corresponding author
    address       : "Leipzig Research Center for Early Child Development, Jahnallee 59, 04109 Leipzig, Germany"
    email         : "manuel.bohn@uni-leipzig.de"
  - name          : "Michael Henry Tessler"
    affiliation   : "3"
  - name          : "Megan Merrick"
    affiliation   : "1"
  - name          : "Michael C. Frank"
    affiliation   : "1"

@@

abstract: |
  Language is learned in complex social settings where listeners must reconstruct speakers’ intentend meanings from context. To navigate this challenge, children can use contextual or pragmatic reasoning to learn the meaning of unfamiliar words. One important challenge for pragmatic reasoning is that it often requires integrating information about the current utterance with broader common ground built up over the course of an interaction. Here we study this integration process. We isolate these two sources of pragmatic information and formalize both how they should be combined and how they might develop using a probabilistic model of conversational reasoning. We present a series of seven experiments with three- to five-year-old children and adults that suggest that these two information sources flexibly trade off with one another. A cue combination model accurately predicted empirical results for children and adults and provided a better fit compared with alternative models in which only one information source was considered. This work integrates distinct sets of findings regarding early language and suggests that pragmatic reasoning models can provide a quantitative framework for understanding developmental changes in language learning.

@@

csl               : pnas.csl

@@

@@

# Significance statement

In order to learn the meaning of a new word, children need to integrate multiple sources of information. Some information is provided by the utterance itself, but other contextual information accumulates over the course of a conversation and is stored in common ground. To understand how these sources are integrated with one another and how they might develop, we formalize this process using computational models of pragmatic reasoning. These models make quantitative predictions for different hypotheses about how integration proceeds. Across seven experiments, we find that the way in which preschoolers and adults integrate information is best described by a model that flexibly trades off between different sources of information. Taken together, this work presents an explicit theory of socially grounded language learning.

@@

What someone means by an utterance is oftentimes not reducible to the words they used. It takes  pragmatic inference – context-sensitive reasoning about the speaker’s intentions - to recover the intended meaning [@grice1991studies, @levinson2000presumptive, @sperber2001relevance]. Contextual information comes in many forms. On the one hand, there is information provided by the utterance\footnote{We use the terms utterance, utterance-level information or utterance-level cues to capture all cues that the speaker provides for their intended meaning. This includes direct referential information in the form of actions such as pointing or gazing, semantic information in the form of conventional word meanings as well as pragmatic inferences that are licenced by the particular choice of words or actions.} itself. Competent language users expect each other to communicate in a cooperative way such that speakers produce utterances that are relevant and informative. Thus, semantic ambiguity can be resolved by reasoning about why the speaker produced this particular utterance [@grice1991studies, @clark1996using, @sperber2001relevance, @tomasello2008origins]. On the other hand, there is information provided by common ground (the body of knowledge and beliefs shared between interlocutors; [@bohn2018common, @clark1996using, @clark2015common]. Because utterances are embedded in a broader (ongoing) social interaction, common ground information serves as a background against which new utterances are interpreted, making some meanings for ambiguous terms more likely than others.

Children learning their first language make inferences about intended meanings based on utterance-level and common-ground information both for language understanding and language learning [@bohn2019pervasive, @clark2009first, @tomasello2008origins]. Starting very early, infants expect adults to produce utterances in a cooperative way [@behne2005one], and expect language to be carrying information [@vouloumanos2012twelve]. By age two, children are sensitive to the informativeness of communication [@o2001two]. By age three children can use this expectation to make pragmatic inferences [@stiller2015ad, @yoon2019role] and to infer novel word meanings [@frank2014inferring]. And although older children continue to struggle with some complex pragmatic inferences until age five and beyond [@noveck2001children], an emerging consensus identifies these difficulties as stemming from difficulties reasoning about linguistic alternatives rather than pragmatic deficits [@skordos2016children, @horowitz2018trouble, @barner2011accessing]. Thus, children’s ability to reason about utterance-level pragmatics is present at least by ages three to five, and possibly substantially younger.

Evidence for the use of common ground information by young children is even stronger: Common ground information guides how infants produce non-verbal gestures and interpret ambiguous utterances [@bohn2018social, @saylor2011s]. For slightly older children, common ground – in the form of knowledge about discourse novelty, preferences, and even discourse expectations – also facilitates word learning [@akhtar1996role, @diesendruck2004two, @saylor2009preschoolers, @sullivan2019discourse].

All of these examples, however, highlight children’s use of a single pragmatic information source or cue. Harnessing multiple – potentially competing – cues poses a separate challenge. One aspect of this integration problem is how to balance common ground information that is built up over the course of an interaction against information gleaned from the current utterance. Much less is known about whether and how children – or even adults – combine these types of information. While many theories of pragmatic reasoning presuppose that both information sources are integrated, the nature of their relationship has typically not been specified. We address this challenge here.

Recent innovations in probabilistic models of pragmatic reasoning provide a quantitative method for addressing the problem of integrating multiple sources of contextual information. This class of computational models, which are referred to as Rational Speech Act (RSA) models [@frank2012predicting, @goodman2016pragmatic] formalize the problem of language understanding as a special case of Bayesian social reasoning. A listener interprets an utterance by assuming it was produced by a cooperative speaker who had the goal to be informative. Being informative is defined as providing a message that would increase the probability of the listener recovering the speaker’s intended meaning in context. This notion of contextual informativeness captures the Gricean idea of cooperation between speaker and listener, and provides a first approximation to what we have described above as utterance-level pragmatic information.

Listeners and speakers also enter into a conversation with assumptions about what is likely to be talked about, a reflection of the common ground shared between them. RSA models capture common ground information as a shared prior distribution over possible intended meanings. Thus, a natural locus for information integration within probabilistic models of pragmatic reasoning is the trade off between the prior probability of a meaning and the informativeness of the utterance. This trade off between contextual factors during word learning is a unique aspect of the word learning problem that is not addressed by other computational models of word learning, which have focused on learning from cross-situational, co-occurrence statistics [@fazly2010probabilistic, @frank2009using] or describing generalizations about word meaning [@xu2007word].

We make use of this framework to study pragmatic cue integration across development. To this end, we adapt a method used in perceptual cue integration studies [@ernst2002humans] predictions about conditions in which they either coincide or conflict.

We start by replicating previous findings with adults showing that listeners make pragmatic inferences based on non-linguistic properties of utterances in isolation (Experiment 1). In separate experiments, we then show that adults make inferences based on common ground information (Experiment 2A and 2B). We use data from these experiments as parameters to generate a priori predictions from RSA models about how utterance and common ground information should be integrated. We consider three models that make different assumptions about the integration process: In the pragmatics model, the two information sources are integrated with one another; according to the flat prior model, participants focus only on the utterance information and in the prior only model, only common ground information is considered. We compare predictions from these models to new empirical data from experiments in which utterance and common ground information are manipulated simultaneously (Experiment 3 and 4).

After successfully validating this approach with adults, we apply the same model-driven experimental procedure to children: We first show that they make pragmatic inferences based on utterance and common ground information separately (Experiment 5 and 6). Then we generate a priori model predictions and compare them to data from an experiment – parallel to Experiment 3 – in which both information sources have to be integrated (Experiment 7). To our knowledge, the strategy of independently estimating model parameters with separate groups and combining them via the model to make preregistered quantitative predictions about a new population has not been used before with children.

Taken together, this work makes two primary contributions: first, it shows that both adults and children integrate utterance-level (Gricean) and common-ground information flexibly. Second, it uses Bayesian data analysis within the RSA framework to provide a model for understanding the multiple loci for developmental change in complex behaviors like contextual communication.

```{r fig1, include = T, fig.align = "center", fig.cap = "Schematic experimental procedure with screenshots from the adult experiments. In all conditions, at test (bottom), the speaker ambiguously requested an object using a non-word (e.g. “dax”). Participants clicked on the object they thought the speaker referred to. Informativeness (Experiment 1, left) translated to making one object less frequent in context. Common ground (Experiment 2, middle) was manipulated by making one object preferred by or new to the speaker. Green plus signs represent utterances that expressed preference and red minus signs represent utterances that expressed dispreference (see main text for details). Experiment 3 (right) combined manipulations. One condition of Experiment 3 is shown here: preference - same speaker - incongruent.", out.width="450px"}
knitr::include_graphics("./figures/fig1.png")
```

# How do adults integrate contextual sources of information?

## Inferences based on utterance and common ground information  (Experiments 1 and 2)

@@

# model
lm_ex1 <- glmer(correct ~ condition +
              (1 |id), 
              data = adult_ex1_data, 
              family = binomial,control=glmerControl(optimizer="bobyqa",optCtrl=list(maxfun=2e5)))

@@

In Experiment 1, participants could learn which object a novel word referred to by assuming that the speaker communicated in an informative way [@frank2014inferring]. The speaker was located between two tables, one with two novel objects, A and B, and the other with only object A (Fig. 1A). When the speaker turned and pointed to the table with the two objects (A and B) and used a novel word to request one of them, participants could infer that the word referred to object B. This follows from the counter-factual inferences that, if the (informative) speaker had wanted to refer to object A, they would have pointed to the table with the single object (this being the least ambiguous way to refer to that object). In the control condition, both tables contained both objects and no inference could be made based on the speaker’s behavior. Participants selected object B above chance in the test condition (t(`r ex1t%>%filter(condition == "test")%>%pull(df)`) = `r ex1t%>%filter(condition == "test")%>%pull(t_value)`, *p* `r ex1t%>%filter(condition == "test")%>%pull(p_value)`) and more often compared to the control condition (*$\beta$* = `r ex1_r%>%filter(term == "conditiontest")%>%pull(estimate)`, se = `r ex1_r%>%filter(term == "conditiontest")%>%pull(std.error)`, *p* `r ex1_r%>%filter(term == "conditiontest")%>%pull(p.value)`). 
```{r ex2-results}
ex2tpref <- adult_ex2_preference_data %>%
  group_by(condition, id) %>%
  summarise(correct = mean(correct)) %>%
  summarise(correct = list(correct)) %>%
  group_by(condition) %>%
  mutate(df= t.test(unlist(correct), mu = 0.5)$parameter,
         t_value = t.test(unlist(correct), mu = 0.5)$statistic,
         p_value = t.test(unlist(correct), mu = 0.5)$p.value) %>%
  select(condition,df,t_value,p_value)%>%
  mutate(p_value = ifelse(p_value<.001,"< .001",as.character(paste("=",substr(round(p_value,3),2,5),sep=" "))),
         t_value = round(t_value,2))

@@

In Experiments 2A and 2B, we tested if participants use common ground information that is specific to a speaker to identify the referent of a novel word [@akhtar1996role, @saylor2009preschoolers]. In Experiment 2A, the speaker expressed a preference for one of two objects (Fig. 1B, left). Later, the speaker used a novel word to request an object. Adults selected the preferred object above chance (t(`r ex2tpref%>%filter(condition == "same_speaker")%>%pull(df)`) = `r ex2tpref%>%filter(condition == "same_speaker")%>%pull(t_value)`, *p* `r ex2tpref%>%filter(condition == "same_speaker")%>%pull(p_value)`) and more so than in a control condition, where a different speaker, whose preferences were unknown, made the request ($\beta$ = `r ex2pref_r%>%filter(term == "conditionsame_speaker")%>%pull(estimate)`, se = `r ex2pref_r%>%filter(term == "conditionsame_speaker")%>%pull(std.error)`, *p* `r ex2pref_r%>%filter(term == "conditionsame_speaker")%>%pull(p.value)`). In Experiment 2B, common ground information came in the form of novelty (Fig 1B, right). First, the speaker encountered one object on one of the tables. Later, a second object appeared. When the same speaker then used a novel word to request an object, participants selected the new object above chance (t(`r ex2tnov%>%filter(condition == "same_speaker")%>%pull(df)`) = `r ex2tnov%>%filter(condition == "same_speaker")%>%pull(t_value)`, *p* `r ex2tnov%>%filter(condition == "same_speaker")%>%pull(p_value)`), and more often compared to when a different speaker (to whom both objects were equally new) made the request ($\beta$ = `r ex2nov_r%>%filter(term == "conditionsame_speaker")%>%pull(estimate)`, se = `r ex2nov_r%>%filter(term == "conditionsame_speaker")%>%pull(std.error)`, *p* `r ex2nov_r%>%filter(term == "conditionsame_speaker")%>%pull(p.value)`). Taken together, Experiments 1 and 2 confirmed that adults make pragmatic inferences based on information provided by the utterance as well as by common ground and provided quantitative estimates of the strength of these inferences for use in our model.

@@

We modeled the integration of utterance informativity and common ground as a process of socially-guided probabilistic inference, using the results of Experiments 1 and 2 to inform key parameters of a computational model. The Rational Speech Act (RSA) model architecture introduced by [@frank2012predicting] encodes conversational reasoning through the perspective of a listener (“he” pronoun) who is trying to decide on the intended meaning of the utterance he heard from the speaker (“she” pronoun). The basic idea is that the listener combines his uncertainty about the speaker’s intended meaning - a prior distribution over referents P(r) - with his generative model of how the utterance was produced: a speaker trying to convey information to him. To adapt this model to the word learning context, we enrich this basic architecture with a mechanism for expressing uncertainty about the meanings of words (lexical uncertainty) - a prior distribution over lexica P(L) [@bergen2016pragmatic].

$$P_{L}(r, \mathcal{L}|u)\propto P_{S}(u|r, \mathcal{L}) \cdot P( \mathcal{L}) \cdot P(r)$$

In the above equation, the listener is trying to jointly resolve the speaker’s intended referent r and the meaning of words (thus learning the lexicon $\mathcal{L}$). He does this by imagining what a rational speaker would say, given the referent they are trying to communicate and a lexicon. The speaker is an approximately rational Bayesian actor (with degree of rationality alpha), who produces utterances as a function of their informativity. The space of utterances the speaker could produce depends upon the lexicon $P(u|\mathcal{L})$; simply put, the speaker labels objects with the true labels under a given lexicon L (see SI Appendix for details):

$$P_{S}(u|r,\mathcal{L})\propto Informativity(u;r)^\alpha \cdot P(u|\mathcal{L})$$

The informativity of an utterance for a referent is taken to be the probability with which a naive listener, who only interprets utterances according to their literal semantics, would select a particular referent given an utterance.

$$Informativity(u; r) = P(r|u) \propto P(r) \cdot \mathcal{L}_{point}$$

The speaker’s possible utterances are pairs of linguistic and non-linguistic signals, namely labels and points. Because the listener does not know the lexicon, the informativity of an utterance comes from the speaker’s point, the meaning of which is encoded in $\mathcal{L}_{point}$ and is simply a truth-function checking whether or not the referent is at the location picked out by the speaker’s point. Though the speaker makes their communicative decision assuming the listener does not know the meaning of the labels, we assume that in addition to a point, the speaker produces a label consistent with their own lexicon $\mathcal{L}$, described by $P(u|\mathcal{L})$ (see SI Appendix for modeling details).

This computational model provides a natural avenue to formalize quantitatively how informativeness and common ground trade-off during word learning. As mentioned above, the common ground shared between speaker and listener plays the role of the listener’s prior distribution over meanings, or types of referents, that the speaker might be referring to and which we posit depends on prior interactions around the referents in the present context (e.g., preference or novelty; Experiment 2A and B). We use the results from Experiment 2 to specify this distribution. The in-the-moment, contextual informativeness of the utterance is captured in the likelihood term, whose value depends on the rationality parameter $\alpha$. Assumptions about rationality may change depending on context and we therefore used the data from Experiment 1 to specify $\alpha$ (see SI Appendix for details about these parameters).
